WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-13 articles/2014/Article-WSClean/wsclean.tex
First draft of article finished.

\begin{abstract}
We present our new imager implementation, ``WSClean'', which we find to be an order of magnitude faster than existing implementations, as well as being capable of MWA full-sky imaging at full resolution.
TODO
\end{abstract}

@@

Observations from non-coplanar interferometric radio telescopes that image large fractions of the sky at once can not be imaged with simple imagers that are based on a two-dimensional Fast Fourier Transformation (FFT). Instead, the imaging algorithm needs to account for the ``w-term'' during inversion, which is the term that describes the deviation of the array from a perfect plane \citep{perley-noncoplanar-arrays}. This deviation is amplified for telescopes with wide field of view, making this especially an issue for low-frequency telescopes, that by nature are wide-field instruments.

There are several methods to deal with the w-term during imaging: facetting \citep{facetting-cornwell}, a three-dimensional Fourier transform \citep{perley-noncoplanar-arrays}, w-projection \citep{wprojection-cornwell}, w-stacking \citep{widefield-imaging-ska-cornwell} and w-snapshots \citep{widefield-imaging-ska-cornwell}. The w-snapshots algorithm is a hybrid between snapshot imaging and one of the other w-term correcting techniques, most efficiently w-projection or w-stacking.

New generation of wide-field observatories are producing data sets that are orders of magnitude larger than before. Examples of such telescopes include the Murchison Widefield Array (MWA, \citealt{mwa}), the upgraded Jansky Very Large Array (JVLA) and the Low-Frequency Array (LOFAR, \citealt{lofar-2013}). The Common Astronomy Software Applications (CASA, \citealt{casa-2007}, \citealt{casa-2008}) has an efficient implementation of the $w$-projection algorithm, with many available features such as multi-scale clean and spectral-shape fitting during deconvolution. However, at the MWA we have seen that imaging a 2-min snapshot observation at low-elevation can take up to tens of hours with CASA's w-projection algorithm. Larger images or higher zenith angles can be impossible to image because the size and number of $w$-kernels become too large to hold in memory.

Another option for reducing MWA data is the Real-Time System (RTS), which has been designed as an efficient calibrating and imaging pipeline specifically for MWA data \citep{rts-mwa}. It can use GPUs for improving efficiency, and uses a w-projection kernel during imaging to correct for the w-term. The kernel can be kept small, because the data is processed in snapshots that have been phase-centered to Zenith. The RTS can not perform deconvolution as of yet.

@@

Once the Square-Kilometre Array (SKA) begins its operation, the required computational power for wide-field imaging will become an even bigger challenge. \citet{widefield-imaging-ska-cornwell} shows that in theory the w-snapshots algorithm is the most efficient approach for the SKA.

@@

This paper is structured as follows: The $w$-stacking algorithm is described in Sect.~\ref{sec:wstacking}. Details of implementing the $w$-stacking and $w$-snapshots algorithms are described in Sect.~\ref{sec:implementation}. The performance and accuracy will be analysed in Sect.~\ref{sec:analysis}.

@@

An interferometer samples the complex visibility function
\begin{align}\notag
V(u,v,w) = & \iint \frac{A(l,m) I(l,m)}{\sqrt{1-l^2-m^2}} \cdot \\ \label{eq:visibility-function}
& e^{-2\pi i \left(ul + vm + w(\sqrt{1-l^2-m^2}-1)\right)} dl dm,
\end{align}
where $u,v,w$ is a baseline coordinate in the coordinate system of the array, $A$ is the primary-beam function, $I$ is the sky function and $l,m$ are cosine sky coordinates. We will use $I'(l,m)$ to denote the sky function before primary-beam correction, $I'(l,m)=A(l,m)I(l,m)$. We will not discuss calibration, but assume $V$ has been calibrated before imaging. In the case of a polarized measurement, the symbols $V$, $A$ and $I$ are $2\times 2$ matrices, but without loss of generality we will ignore polarization and treat inversion as a scalar problem. Imaging consists of inverting Eq.~\eqref{eq:visibility-function}, i.e., to find $I'$ from $V$.

For small field of views, the term $\sqrt{1-l^2-m^2}$ is approximately of unit size, making Eq.~\eqref{eq:visibility-function} approximate an ordinary 2-dimensional Fourier transform, which can be inverted with an inverse Fast Fourier Transform (FFT). A common rule is that this is valid when
\begin{equation}\label{eq:when-2d-is-valid}
\forall w,l,m: w\left(\sqrt{1-l^2-m^2}-1\right) \ll 1.
\end{equation}

To derive the w-stacking technique, Eq.~\eqref{eq:visibility-function} is rewritten to
\begin{align}\notag
V(u,v,w) = & \iint \frac{I'(l,m) e^{-2\pi i w(\sqrt{1-l^2-m^2}-1)}}{\sqrt{1-l^2-m^2}} \cdot \\ \notag
& e^{-2\pi i \left(ul + vm\right)} dl dm.
\end{align}
This is an ordinary two-dimensional Fourier transform going from $u,v$ space to $l,m$ space, and can be inverted to get:
% Intermediate step:
%\begin{align}\notag
%\frac{I'(l,m) e^{-2\pi i w(\sqrt{1-l^2-m^2}-1)}}{\sqrt{1-l^2-m^2}} = & \iint V(u,v,w) \cdot \\ \notag
%& e^{-2\pi i \left(ul + vm\right)} du dv,
%\end{align}
%which can be rewritten to
\begin{align}\notag
\frac{I'(l,m)}{\sqrt{1-l^2-m^2}} = & e^{2\pi i w(\sqrt{1-l^2-m^2}-1)} \iint V(u,v,w) \cdot \\ \notag
& e^{2\pi i \left(ul + vm\right)} du dv.
\end{align}
Integrating both sides over $w_0$ to $w_e$, the minimum and maximum value of $w$, results in
\begin{align}\notag
\frac{I'(l,m)\left(w_e - w_0\right)}{\sqrt{1-l^2-m^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-l^2-m^2}-1)} \cdot \\ \label{eq:wstacking}
\iint V(u,v,w)  e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
The final step is to make the $u,v,w$ parameters discrete, so that the integration over $u$ and $v$ can become an inverse FFT and the integration over $w$ becomes a summation. This shows that the sky function can be reconstructed by i)~grid samples with equal $w$-value on a uniform grid; ii)~calculating the inverse FFT; iii)~apply the direction dependent phaseshift $e^{2\pi i w(\sqrt{1-l^2-m^2}-1)}$; iv)~repeat this for all $w$-values and add the results together; v)~apply the final scaling.

In practice, the final scaling will be different from $\left(w_e - w_0\right)/\sqrt{1-l^2-m^2}$ suggested by Eq.~\eqref{eq:wstacking}, because the individual $w$-layers will not be completely filled with samples. Therefore, the number of added samples and their weight have to be taken into account. Additionally, it might be required to divide out the effect of a possible convolution kernel, the primary beam and correct for other direction dependent effects. In XX, YY, LL or RR polarizations, a correlated baseline is the complex conjugate of the reversed baseline, and the relation $V(u,v,w)=\overline{V(-u,-v,-w)}$ holds. The right-hand side of Eq.\eqref{eq:wstacking} with only positive $w$-value samples then becomes the complex conjugate of the one with only negative $w$-values samples. In this case we can therefore calculate the image for $w<0$ from the image with $w>0$. That allows us to set $w_0$ to the minimum absolute $w$-value, which requires at least two times less layers. In any case, the input to the two-dimensional inverse FFT is not generally an Hermitian-symmetric function, and hence is always performed as a complex-to-complex transform.

@@

Using too few $w$-layers results in decorrelation of off-axis sources in the longer baselines. Additionally, $w$-aliasing errors can cause ghost sources in the image. An (extreme) example of this effect is shown in Fig.~\ref{fig:aliasing-example}. When Cotton-Schwab cleaning includes prediction with too few $w$-layers, incorrect values will be subtracted from the visibilities even when no aliased sources are cleaned during minor iterations. Therefore, accurate prediction is more important than accurate imaging, because aliasing artefacts are attenuated by cleaning as long as the model is subtracted accurately. Possible ways to allow less $w$-layers during imaging is by predicting with more $w$-layers or a direct Fourier transform. We have not looked into these options yet.

@@

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by w-stacking or by w-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_e - w_0\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \cdot \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. By doing the inverse corrections during prediction, such a shifted visibility set can be cleaned with Cotton-Schwab iterations similar to a non-shifted set.

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. A recentred image can be stored in a Flexible Image Transport System (FITS) file, which allows setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. An example of the difference in projection is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13~min MWA observation imaged with WSClean. Cleaning an image that is in a different projection yields slightly different results, but qualitativily the two images are clearly of equal accuracy. The wall-clock time for imaging is 60 min and 41 min for respectively normal projection and the recentred image. Of the 41 min, 3 min is spent on phase rotating the visibilities.

@@

The used software version of CASA is ``stable release 42.0, revision 26465'', which was released September 2013. It includes performance improvements to the \texttt{clean} task which have recently been made. For WSClean, the latest version as of February 2014 was used. The tests were run on a high-end desktop with 32 GB of memory and a 3.20-GHz Intel Core i7-3930K processor with 6 cores that can perform 138 giga-floating point operations per second (GFLOPS). The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s.
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & XX \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

@@

Table~\ref{tbl:accuracy-measurements} lists the measured RMS in the residual image and the standard errors of the source brightnesses as detected by Aegean. In all tests, WSClean performs slightly more accurate. At zenith, CASA produces a residual RMS twice as high as WSClean. This is caused by the fact that we keep the number of $w$-projection planes in w-projection equal to the number of w-layers in w-stacking, resulting in only 12 $w$-planes at zenith. This evidently has a stronger effect on the $w$-projection algorithm. However, the computational performance of the $w$-projection algorithm is hardly affected by the number of $w$-projection planes, and in practical situations one would always use more w-projection planes\footnote{In the zenith case CASA issues a warning to use more $w$-projection planes.}. When 128 $w$-projection planes are used in CASA, its residual RMS is equal to WSClean with 12 $w$-layers, but the flux density measurements are less accurate. We do not know why this parameter needs to be higher in CASA to reach the same RMS.\footnote{TODO: do I want to find out this is because of invalid sqrt(w) gridding?} Eq.~\eqref{eq:nwlayers-bound} results in $N_\textrm{wlayers}\gg 6$ for the top-left image pixel that is 26\degree away from the centre, while the sources are all within 10\degree. Also unexpected is that the source flux density becomes worse by increasing the number of planes. For WSClean both values stay approximately the same when the number of $w$-layers is increased.

In the $\textrm{ZA}=10\degree$ case, CASA is slightly less accurate. The extra noise can be observed in the produced images, as shown in Fig.~\ref{fig:residuals}. Also made is an image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}. Its accuracy is higher compared to the normal projection, but its residual noise is also higher. The reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

WSClean is faster in all tested cases. WSClean and CASA determine a different next major iteration threshold. CASA performs 5 major iterations at zenith and 8 at 10\degree angle, while WSClean always performs 5 major iterations. Major iterations are dominating the computing time. We will look more closely at the difference in performance in the next section.

@@

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_e,N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_e + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_e,N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_e + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_e$ is the maximum $w$ value and is estimated with
\begin{equation}
 w_e = \frac{1}{\lambda}\left[\max |\Delta \mathbf{x}| \sin \textrm{ZA} + \max \Delta z \cos \textrm{ZA} \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right],
\end{equation}
where $\max |\Delta \mathbf{x}|$ is the maximum baseline length, $\max \Delta z$ is the maximum height difference between antennas and $\lambda$ is the wavelength. These functions follow the trend of the measurements well, with an absolute error of 14.7\% and 20.7\% for respectively the WSClean and CASA functions.

@@

Function $t_\textrm{snapshot}$ can be minimized to find the optimal snapshot duration. For WSClean with MWA parameters, we find that this function decreases but no minimum is reached for $\Delta \tau_\textrm{snapshot}<24$~h. Therefore, from a performance perspective, the snapshot duration should be as large as possible. For $w$-projection in CASA we find an optimal snapshot duration of 122~s for MWA observations.

\begin{table*}
\caption{Configurations for which the computational cost are predicted.} \label{tbl:computational-cost-configurations}
\begin{tabular}{l|rrrrrrrrrr}
 Configuration&$\lambda$& FOV  & Beams & Ant & Res & $\max |\Delta \mathbf{x}|$ & $\max \Delta z$ & int. t & BW & $\Delta \nu$ \\
          & (m)       & (FWHM)&       &     &     & (km)                       & (m)             & (s)    & (MHz) &  (kHz) \\
\hline
 GLEAM    &  2   & 24.7\degree& 1    & 128 & 2'   & 2.9 & 5 & 2 & 32 & 40 \\
 EMU      & 0.2  & 1\degree   & 30   &  36 & 10''  & 6 &  0 &10 & 300 & 20\\
 MSSS low &  5   & 9.8\degree & 5    &  20 & 100'' & 5 &    &10 & 16 & 16 \\
 MSSS high&  2   & 3.8\degree & 5    &  40 & 120'' & 5 &    &10 & 16 & 16 \\
 AARTFAAC &  5   & 45\degree  & 1    & 288 &  20'  & 0.3& 1 & 1 &  7 & 24 \\
 VLSS     & 4.1  & 14\degree  & 1    &  27 &  80'' &11.1& 0 &10 & 1.56 & 12.2 \\
 SKA-P1 AA inner&3& 1.2\degree& 480  &  35 & 140'' &2.5& 2 &  8 & 380 & 80 \\
 SKA-P1 AA +mid&3 & 1.2\degree & 480  &  50 & 3.5'' &100& 50 & 0.2 & 380 & 2 \\
\end{tabular}
\end{table*}

>>>>

\begin{abstract}
We present a new imager implementation that uses the $w$-stacking algorithm and can make use of the $w$-snapshot algorithm. On MWA data, we find it to be an order of magnitude faster than $w$-projection, as well as being capable of full-sky imaging at full resolution. The performance dependencies of CASA's $w$-projection and WSClean are analysed and analytical functions are derived that describe the required computing cost for both imagers. The difference in computing cost is mainly caused by the efficiency of our new imager on MWA's 27\degree field of view. We predict the computing costs for several other configurations and estimate that our imager is approximately a factor of 2--3 faster on other arrays. Especially imaging SKA observations will become challenging, for which we predict that our new algorithm is still at least 30 times slower than is budgetted. We find that the $w$-snapshot has only a minor benefit. The source code of our new imager is publicly released.
\end{abstract}

@@

Observations from non-coplanar interferometric radio telescopes that image large fractions of the sky at once can not be imaged with a two-dimensional Fast Fourier Transformation (FFT). Instead, the imaging algorithm needs to account for the ``w-term'' during inversion, which is the term that describes the deviation of the array from a perfect plane \citep{perley-noncoplanar-arrays}. This deviation is amplified for telescopes with wide field of view, making this especially an issue for low-frequency telescopes that by nature are wide-field instruments.

There are several methods to deal with the w-term during imaging: facetting \citep{facetting-cornwell}, a three-dimensional Fourier transform \citep{perley-noncoplanar-arrays}, w-projection \citep{wprojection-cornwell}, w-stacking \citep{widefield-imaging-ska-cornwell} and w-snapshots \citep{widefield-imaging-ska-cornwell}.

New generation of wide-field observatories are producing data sets that are orders of magnitude larger than before. Examples of such telescopes include the Murchison Widefield Array (MWA, \citealt{mwa}), the upgraded Jansky Very Large Array (JVLA) and the Low-Frequency Array (LOFAR, \citealt{lofar-2013}). The Common Astronomy Software Applications (CASA, \citealt{casa-2007}, \citealt{casa-2008}) have an efficient implementation of the $w$-projection algorithm, with many available features such as multi-scale clean and spectral-shape fitting during deconvolution. However, at the MWA we have seen that imaging a 2-min snapshot observation at low-elevation can take up to tens of hours with CASA's w-projection algorithm. Larger images or higher zenith angles can be impossible to image because the size and number of $w$-kernels become too large to hold in memory.

Another option exists for imaging MWA data: the Real-Time System (RTS, \citealt{rts-mwa}). This has been designed as an efficient calibration and imaging pipeline specifically for MWA data. It can use GPUs for improving efficiency, and uses a w-projection kernel during imaging to correct for the w-term. The kernel can be kept small, because the data is processed in snapshots that have been phase-centered to zenith. The RTS can not perform deconvolution as of yet.

@@

Once the Square-Kilometre Array (SKA) begins its operation, the required computational power for wide-field imaging will become an even bigger challenge. \citet{widefield-imaging-ska-cornwell} argues that the w-snapshots algorithm is the most efficient approach for the SKA.

@@

This paper is structured as follows: The $w$-stacking algorithm is described in Sect.~\ref{sec:wstacking}. Details of implementing the $w$-stacking and $w$-snapshots algorithms are described in Sect.~\ref{sec:implementation}. The performance and accuracy will be analysed in Sect.~\ref{sec:analysis}. The conclusions are presented in Sect.~\ref{sec:conclusions}.

@@

An interferometer samples the complex visibility function
\begin{align}\notag
V(u,v,w) = & \iint \frac{A(l,m) I(l,m)}{\sqrt{1-l^2-m^2}} \times \\ \label{eq:visibility-function}
& e^{-2\pi i \left(ul + vm + w(\sqrt{1-l^2-m^2}-1)\right)} dl dm,
\end{align}
where $u,v,w$ is a baseline coordinate in the coordinate system of the array, $A$ is the primary-beam function, $I$ is the sky function and $l,m$ are cosine sky coordinates. We will use $I'(l,m)$ to denote the sky function before primary-beam correction, $I'(l,m)=A(l,m)I(l,m)$. We will not discuss calibration, but assume $V$ has been calibrated before imaging. In the case of a polarized measurement, the symbols become $2\times 2$ matrices, but without loss of generality we will ignore polarization and treat inversion as a scalar problem. Imaging consists of inverting Eq.~\eqref{eq:visibility-function}, i.e., to find $I'$ from $V$.

For small field of views, the term $\sqrt{1-l^2-m^2}$ is approximately of unit size, making Eq.~\eqref{eq:visibility-function} approximate an ordinary 2-dimensional Fourier transform that can be inverted with an inverse Fast Fourier Transform (FFT). A common rule is that this is valid when
\begin{equation}\label{eq:when-2d-is-valid}
\forall w,l,m: w\left(\sqrt{1-l^2-m^2}-1\right) \ll 1.
\end{equation}

To derive the w-stacking technique, Eq.~\eqref{eq:visibility-function} is rewritten to
\begin{align}\notag
V(u,v,w) = & \iint \frac{I'(l,m) e^{-2\pi i w(\sqrt{1-l^2-m^2}-1)}}{\sqrt{1-l^2-m^2}} \times \\ \notag
& e^{-2\pi i \left(ul + vm\right)} dl dm.
\end{align}
This is an ordinary two-dimensional Fourier transform going from $u,v$ space to $l,m$ space, and can be inverted to get:
% Intermediate step:
%\begin{align}\notag
%\frac{I'(l,m) e^{-2\pi i w(\sqrt{1-l^2-m^2}-1)}}{\sqrt{1-l^2-m^2}} = & \iint V(u,v,w) \times \\ \notag
%& e^{-2\pi i \left(ul + vm\right)} du dv,
%\end{align}
%which can be rewritten to
\begin{align}\notag
\frac{I'(l,m)}{\sqrt{1-l^2-m^2}} = & e^{2\pi i w(\sqrt{1-l^2-m^2}-1)} \iint V(u,v,w) \times \\ \notag
& e^{2\pi i \left(ul + vm\right)} du dv.
\end{align}
Integrating both sides over $w_0$ to $w_e$, the minimum and maximum value of $w$, results in
\begin{align}\notag
\frac{I'(l,m)\left(w_e - w_0\right)}{\sqrt{1-l^2-m^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-l^2-m^2}-1)} \times \\ \label{eq:wstacking}
\iint V(u,v,w)  e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
The final step is to make the $u,v,w$ parameters discrete, so that the integration over $u$ and $v$ can become an inverse FFT and the integration over $w$ becomes a summation. This shows that the sky function can be reconstructed by i)~grid samples with equal $w$-value on a uniform grid; ii)~calculating the inverse FFT; iii)~apply the direction dependent phaseshift $e^{2\pi i w(\sqrt{1-l^2-m^2}-1)}$; iv)~repeat this for all $w$-values and add the results together; v)~apply the final scaling.

In practice, the final scaling will be different from $\left(w_e - w_0\right)/\sqrt{1-l^2-m^2}$ suggested by Eq.~\eqref{eq:wstacking}, because the individual $w$-layers will not be completely filled with samples. Therefore, the number of added samples and their weight have to be taken into account. Additionally, it might be required to divide out the effect of a possible convolution kernel, the primary beam and correct for other direction dependent effects. In XX, YY, LL or RR polarizations, a correlated baseline is the complex conjugate of the reversed baseline, and the relation $V(u,v,w)=\overline{V(-u,-v,-w)}$ holds. The right-hand side of Eq.\eqref{eq:wstacking} with only positive $w$-value samples then becomes the complex conjugate of the one with only negative $w$-values samples. In this case we can therefore calculate the image for $w<0$ from the image with $w>0$. That allows us to set $w_0$ to the minimum absolute $w$-value, which requires two times fewer layers. In any case, the input to the two-dimensional inverse FFT is not generally an Hermitian-symmetric function, and hence the inverse FFT is always performed as a complex-to-complex transform.

@@

Using too few $w$-layers results in decorrelation of off-axis sources in the longer baselines. Additionally, $w$-aliasing errors can cause ghost sources in the image. An (extreme) example of this effect is shown in Fig.~\ref{fig:aliasing-example}. When Cotton-Schwab cleaning includes prediction with too few $w$-layers, incorrect values will be subtracted from the visibilities even when no aliased sources are cleaned during minor iterations. Therefore, accurate prediction is more important than accurate imaging, because aliasing artefacts are attenuated by cleaning as long as the model is subtracted accurately.

@@

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by w-stacking or by w-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_e - w_0\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \times \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. By doing the inverse corrections during prediction, such a shifted visibility set can be cleaned with Cotton-Schwab iterations similar to a non-shifted set.

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. A recentred image can be stored in a Flexible Image Transport System (FITS) file with the slant orthographic (SIN) projection normally used in synthesis imaging, by setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. An example of the difference in projection is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13~min MWA observation imaged with WSClean. Cleaning an image that is in a different projection yields slightly different results, but qualitativily the two images are clearly of equal accuracy. The wall-clock time for imaging is 60 min and 41 min for respectively normal projection and the recentred image. Of the 41 min, 3 min is spent on phase rotating the visibilities.

@@

The used software version of CASA is ``stable release 42.0, revision 26465'', which was released September 2013. For WSClean, the latest version as of February 2014 was used. The tests were run on a high-end desktop with 32 GB of memory and a 3.20-GHz Intel Core i7-3930K processor with 6 cores that can perform 138 giga-floating point operations per second (GFLOPS). The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s.
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & XX \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

@@

Table~\ref{tbl:accuracy-measurements} lists the measured RMS in the residual image and the standard errors of the source brightnesses as detected by Aegean. In all tests, WSClean performs slightly more accurate. At zenith, CASA produces a residual RMS twice as high as WSClean. This is caused by the fact that we keep the number of $w$-projection planes in w-projection equal to the number of w-layers in w-stacking, resulting in only 12 $w$-planes at zenith. This evidently has a stronger effect on the $w$-projection algorithm. However, the computational performance of the $w$-projection algorithm is hardly affected by the number of $w$-projection planes, and in practical situations one would always use more w-projection planes. When 128 $w$-projection planes are used in CASA, its residual RMS is equal to WSClean with 12 $w$-layers, but the flux density measurements are less accurate. We do not know why this parameter needs to be higher in CASA to reach the same RMS (TODO: do I want to find out this is because of invalid sqrt(w) gridding?).  Eq.~\eqref{eq:nwlayers-bound} results in $N_\textrm{wlayers}\gg 6$ for the top-left image pixel that is 26\degree away from the centre, while the sources are all within 10\degree. Also unexpected is that the source flux density becomes worse by increasing the number of planes. For WSClean both values stay approximately the same when the number of $w$-layers is increased.

In the $\textrm{ZA}=10\degree$ case, CASA is slightly less accurate. The extra noise can be observed in the produced images, as shown in Fig.~\ref{fig:residuals}. An image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}, is also made and analyzed. Its accuracy is higher compared to the normal projection, but its residual noise is also higher. The reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

WSClean is faster in all tested cases. Both imagers have performed 5 major iterations for these results, and major iterations are dominating the computing time. We will look more closely at the difference in performance in the next section.

@@

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_e,N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_e + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_e,N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_e + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_e$ is the maximum $w$ value and is estimated with
\begin{equation}
 w_e = \frac{1}{\lambda}\left[\max |\Delta \mathbf{x}| \sin \textrm{ZA} + \max \Delta z \cos \textrm{ZA} \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right],
\end{equation}
where $\max |\Delta \mathbf{x}|$ is the maximum baseline length, $\max \Delta z$ is the maximum height difference between antennas and $\lambda$ is the wavelength. These functions follow the trend of the measurements well, with an absolute error of 14.7\% and 20.7\% for respectively the WSClean and CASA functions. (TODO $w_e$ is slightly different for CASA)

@@

Function $t_\textrm{snapshot}$ can be minimized to find the optimal snapshot duration. For WSClean with MWA parameters, we find that this function decreases but no minimum is reached for $\Delta \tau_\textrm{snapshot}<24$~h. Therefore, from a performance perspective, the snapshot duration should be as large as possible. For $w$-projection in CASA we find an optimal snapshot duration of $\sim2$ min for MWA observations.

\begin{table*}
\caption{Configurations for which the computational cost are predicted.} \label{tbl:computational-cost-configurations}
\begin{tabular}{l|rrrrrrrrrr}
 Configuration&$\lambda$& FOV  & Beams & Ant & Res & $\max |\Delta \mathbf{x}|$ & $\max \Delta z$ & int. t & BW & $\Delta \nu$ \\
          & (m)       & (FWHM)&       &     &     & (km)                       & (m)             & (s)    & (MHz) &  (kHz) \\
\hline
 GLEAM    &  2   & 24.7\degree& 1    & 128 & 2'   & 2.9 & 5 & 2 & 32 & 40 \\
 EMU      & 0.2  & 1\degree   & 30   &  36 & 10''  & 6 & 0.2 &10 & 300 & 20\\
 MSSS low &  5   & 9.8\degree & 5    &  20 & 100'' & 5 & 2  &10 & 16 & 16 \\
 MSSS high&  2   & 3.8\degree & 5    &  40 & 120'' & 5 & 2  &10 & 16 & 16 \\
 LOFAR LBA NL& 5 & 4.9\degree & 1    &  38 &   3'' &180& 20 & 1 & 96 & 1 \\
 AARTFAAC &  5   & 45\degree  & 1    & 288 &  20'  & 0.3& 0.5 & 1 &  7 & 24 \\
 VLSS     & 4.1  & 14\degree  & 1    &  27 &  80'' &11.1& 0.2 &10 & 1.56 & 12.2 \\
 MeerKAT  & 0.2  &  1\degree  & 1    &  64 &   6'' &  8 & 1   & 0.5 & 750  & 50 \\
 SKA-P1 AA inner&3& 1.2\degree& 480  &  35 & 140'' &2.5& 2 &  8 & 380 & 80 \\
 SKA-P1 AA +mid&3 & 1.2\degree & 480  &  50 & 3.5'' &100& 50 & 0.2 & 380 & 2 \\
\end{tabular}
\end{table*}
